Spontaneous Similarity Judgments of Musical Tones played with various Playing Techniques

### 2018-01-15 paper/lagrangeSpontaneousSimilarity2017.tex
good dendrogram

\author{Christian El-Hajj, Mathias Rossignol, Grégoire Lafay, Mathieu Lagrange}

@@

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/clusterAnalysis.png}
\caption{Cluster analysis of the similarity judgments. $j/I$ and $j/Pt$ compare the partition obtained using the similarity judgments clustered using a growing number of clusters versus respectively the instrument partition and the playing technique partition. Normalizing null partitions based NMI $null$ is used to obtain normalized measures: $nj/I$ and $nj/Pt$. \ml{change legend, and use only one null}}
\label{fig:clusters}
\end{figure}

There is a clear gain of NMI when considering the playing technique which means that the subjects primarily used cues corresponding to the playing technique to organize the \ipts. Also, for both curves, a global trend of increase in the $j/x$ partitions with respect to the number of clusters can be observed. This is a well documented fact, which simply put, means that it is easier to match any given partition with more clusters at hand \cite{tibshirani2001estimating}. To compensate for this bias, one can consider the NMI that would be achieved by comparing to the reference partition several random partitions of increasing number of clusters. This so-called null partition is termed $null$. The resulting averaged NMI for 100 randomly generated partitions is displayed on Figure \ref{fig:clusters}. Normalizing $j/I$ and $j/Pt$ to reduce the impact of this bias, we simply substract the $null$ NMI:
\begin{eqnarray}
  nj/I &=& j/I - null \\
  nj/Pt &=& j/Pt - null  \\
\end{eqnarray}
This normalization is useful for identifying more easily the number of clusters for which the comparison is most relevant. As can be seen on Figure \ref{fig:clusters}, the number of clusters that leads to the highest NMI is 5 for the playing technique reference and 8 for the instrument reference, leading respectively to an NMI of $0.46$ and $0.21$. For the latter, reducing the number of clusters down to 6 do not decrease the NMI by much, a fact that will be latter discussed in more details. Even by considering the null normalization, the matching level is much higher for the playing technique reference.

>>>>

\author{Christian El-Hajj, Mathias Rossignol, Gr\'egoire Lafay, Mathieu Lagrange}

@@

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/clusterAnalysis.png}
\caption{Cluster analysis of the similarity judgments. $j/I$ and $j/Pt$ compare the partition obtained using the similarity judgments clustered using a growing number of clusters versus respectively the instrument partition and the playing technique partition. Normalizing null partitions based NMI $nullI$ and $nullPt$ are used to obtain normalized measures: $nj/I$ and $nj/Pt$.}
\label{fig:clusters}
\end{figure}

There is a clear gain of NMI when considering the playing technique which means that the subjects primarily used cues corresponding to the playing technique to organize the \ipts. Also, for both curves, a global trend of increase in the $j/x$ partitions with respect to the number of clusters can be observed. This is a well documented fact, which simply put, means that it is easier to match any given partition with more clusters at hand \cite{tibshirani2001estimating}. To compensate for this bias, one can consider the NMI that would be achieved by comparing to the reference partition several random partitions of increasing number of clusters. The resulting averaged NMI for 100 randomly generated partitions is displayed on Figure \ref{fig:clusters}. This so-called null partitions NMI are respectively termed $nullI$ and $nullPt$ depending on the reference used. To normalize $j/I$ and $j/Pt$ in order to reduce the impact of this bias, we simply substract the $null$ NMI:
\begin{eqnarray}
  nj/I &=& j/I - null \\
  nj/Pt &=& j/Pt - null  \\
\end{eqnarray}
This normalization is useful for identifying more easily the number of clusters for which the comparison is most relevant. As can be seen on Figure \ref{fig:clusters}, the number of clusters that leads to the highest NMI is 5 for the playing technique reference and 8 for the instrument reference, leading respectively to an NMI of $0.46$ and $0.21$. For the latter, reducing the number of clusters down to 6 do not decrease the NMI by much, a fact that will be latter discussed in more details. Even by considering the null normalization, the matching level is much higher for the playing technique reference.
